---
title: "Data Replication Assignment using Gelfand et al. 2021"
author: "Ajna F. Kertesz"
date: "4/13/2022"
output:
  html_document:
    toc: yes
    theme: united
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "images/")
```

## Briefly about the project

This project is aimed at replicating the major findings of Gelfand et al. 2021 using open access data on COVID-19, as well as CTL scores by countries. 

The original cross-national analysis used Cultural Tightness Looseness (CTL) Scores to predict COVID-19 cases and death. Previous studies indicate that there is a large between-country variation in CTL scores. Some countries are more "tight", which means they have more norms in their society and they follow these norms more closely, while other countries are more "loose", with less norms and expectations in their society. The COVID-19 pandemic presented an interesting situation, where norm conformity and compliance with new norms and health recommendations (e.g., mask-wearing, social distancing...etc) became a direct predictor of infection and death rates. Therefore, Gelfand and her colleges were interested in investigating whether traditionally tighter countries did indeed have lower infection and death rates. 

First they ran simple correlations between centered CTL scores and infection and death rates. They graphed the line of best fit for these two correlation.
Next, they used ordinary least squares regression models where CTL was used as a predictor of COVID-19 infection rates. After this, they ran a second regression using CTL as a predictor of COVID-19 death rates. In both models, they adjusted for a number of covariaties including GINI, population density, power distance, local stay-at-home mandates...etc.

Below, I am replicating these models and plots.


```{r lib and data, echo=FALSE, message=FALSE, warning=FALSE}
# load all libraries that will be needed
library(dbplyr)
library(tidyr)
library(readr)
library(tidyverse)
library(car)
library(ggplot2)
library(ggrepel)
library(sjPlot)

#read in data

data <- read_csv("data/Data.csv")
```


```{r explore and transfrom data, echo=FALSE}



#centering variables by zscorsing them using the scale function

data<-data%>%
  mutate(ZTightness        = scale(Tightness),
         ZStringencyIndex  = scale(StringencyIndex),
         Zday_diff         = scale(day_diff),
         ZPowerdistance    = scale(Powerdistance),
         ZRelatiolMobility = scale (RelatiolMobility),
         ZGovernment.efficiency = scale (Government.efficiency),
         ZGDP.capita = scale(GDP.capita),
         ZGini = scale(Gini),
         ZMedian_Age = scale(Median_Age),
         ZUnderreporting = scale(Underreporting),
         ZAuthoritarian = scale(Authoritarian),
         ZMortality = scale(Mortality), 
         ZpercentMigrants = scale(percentMigrants), 
         ZOct16_total_cases_per_million = scale(Oct16_total_cases_per_million),
         ZOct16_total_deaths_per_million= scale(Oct16_total_deaths_per_million),
         ZOct16_total_tests_per_thousand= scale(Oct16_total_tests_per_thousand))


#calculating and centering collectivism based on individualism
data$Collectivism<-(0-data$Individualism)+100
data$ZCollectivism<-scale(data$Collectivism)


#using log calculation to decrease skew in these large datasets with potentially many outliers (e.g. depending on healthcare quality and access these numbers could look vastly different)
data$logpopulation<-log(data$population)
data$Zlog_PopDensity<-scale(log(data$Population.density+1))
data$ZLogBeds<-scale(log(data$hospital_beds_per_thousand))
data$OCTCases<-log(data$Oct16_total_cases_per_million)
data$OCTDeaths<-log(data$Oct16_total_deaths_per_million)

#calculating a distribution of test/cases (in many countries only the really sick got tested, which might make it seem like the cases were low)
data$TestsPerCaseOct<-data$Oct16_total_tests/data$Oct16_total_cases
data$ZTestsPerCaseOct<-scale(data$TestsPerCaseOct)


```

## Reproducing Figue 1

The association of cultural tightness and logged cases per million (Oct 2020) *the scatterplot doesn't include any covariates


```{r cor 1, echo=FALSE, message=FALSE, warning=FALSE}

#run correlation test for tightness and cases without any controls
cor.test(data$ZTightness,data$OCTCases) #p=-0.406

figure1<-ggplot(data, aes(x = Tightness, y = OCTCases))+
  geom_point(colour = "red", size=1) + 
  geom_text(hjust=0.005, nudge_x = 0.005, label=data$Country, size=1.5, check_overlap = TRUE)+
geom_smooth(method = "lm", fill = NA, se=TRUE)+
  labs(x="Cultural Tightness", y="Log of COVID-19 cases per million as of Oct 16")+
  theme(legend.position = "none")

figure1

# I tried running their original code, but it doesn't include the labels for the countries
#plot1<-ggplot(data, aes(x=Tightness, y=OCTCases))+
#  labs(x="Cultural Tightness",y="Log of COVID-19 Cases Per Million as of October 16th")+
#  theme_classic(base_size = 15)+
#  geom_smooth(method=lm, se=TRUE, color="black", fill = "#B8B8B8")+
#  geom_point(color="#cd5b45", size=3)+
# theme(legend.position = "none",axis.text.x = element_text(color="#000000"),
#        axis.text.y = element_text(color="#000000")) 
# plot1
 
#save Figure 1
ggsave("cases_oct.png", figure1,width=12,height=12,units="in")


```

## Reproducing Figue 2

The association of cultural tightness and logged deaths per million (Oct 2020) *the scatterplot doesn't inlcude any covariates

```{r cor 2, echo=FALSE, message=FALSE, warning=FALSE}


#run corrolation test for tightness and death without any controls
cor.test(data$ZTightness,data$OCTDeaths) #p=-0.4537

#make ggplot of the correlation
figure2<-ggplot(data, aes(x = Tightness, y = OCTDeaths))+
  geom_point(colour = "red", size=1) + 
  geom_text(hjust=0.005, nudge_x = 0.005, label=data$Country, size=2, check_overlap = TRUE)+
geom_smooth(method = "lm", fill = NA, se=TRUE)+
  labs(x="Cultural Tightness", y="Log of COVID-19 cases per million as of Oct 16")+
  theme(legend.position = "none")

figure2

#save Figure 2
ggsave("deaths_oct.png", figure2,width=12,height=12,units="in")

```

## Reproducing Table 1 

### Logged Ñases per million (10 regression models with different covariates)

```{r table1, echo=FALSE}

#creating the regressions with the variables included in table 1.
#using the lm function and summarizing them by using summary

summary(Model1<- lm((OCTCases) ~ 
                 ZTestsPerCaseOct+
                 ZGDP.capita+
                 ZGini+
                 Zlog_PopDensity+
                 ZpercentMigrants+
                 ZGovernment.efficiency,data=data))

summary(Model2<- lm((OCTCases) ~ 
                 ZTestsPerCaseOct+
                 ZGDP.capita+
                 ZGini+
                 Zlog_PopDensity+
                 ZpercentMigrants+
                 ZGovernment.efficiency+
                 ZTightness, data=data))

summary(Model3<- lm((OCTCases) ~ 
                 ZOct16_total_tests_per_thousand+
                 ZGDP.capita+
                 ZGini+
                 Zlog_PopDensity+
                 ZpercentMigrants+
                 ZGovernment.efficiency+
                 ZTightness, data=data))

summary(Model4<- lm((OCTCases) ~ 
                 ZUnderreporting+
                 ZGDP.capita+
                 ZGini+
                 Zlog_PopDensity+
                 ZpercentMigrants+
                 ZGovernment.efficiency+
                 ZTightness, data=data))

summary(Model5<- lm((OCTCases) ~ 
                 ZLogBeds+
                 ZGDP.capita+
                 ZGini+
                 Zlog_PopDensity+
                 ZpercentMigrants+
                 ZGovernment.efficiency+
                 ZTightness, data=data))

summary(Model6<- lm((OCTCases) ~ 
                 ZTestsPerCaseOct+
                 ZGDP.capita+
                 ZGini+
                 Zlog_PopDensity+
                 ZpercentMigrants+
                 ZGovernment.efficiency+
                 ZTightness+
                 ZCollectivism+
                 ZPowerdistance, data=data))

summary(Model7<- lm((OCTCases) ~ 
                 ZTestsPerCaseOct+
                 ZGDP.capita+
                 ZGini+
                 Zlog_PopDensity+
                 ZpercentMigrants+
                 ZGovernment.efficiency+
                 ZTightness+
                 ZAuthoritarian, data=data))

summary(Model8<- lm((OCTCases) ~ 
                 ZTestsPerCaseOct+
                 ZGDP.capita+
                 ZGini+
                 Zlog_PopDensity+
                 ZpercentMigrants+
                 ZGovernment.efficiency+
                 ZTightness+
                 ZMedian_Age, data=data))

summary(Model9<- lm((OCTCases) ~ 
                 ZTestsPerCaseOct+
                 ZGDP.capita+
                 ZGini+
                 Zlog_PopDensity+
                 ZpercentMigrants+
                 ZGovernment.efficiency+
                 ZTightness+
                 Zday_diff, data=data))

summary(Model10<- lm((OCTCases) ~ 
                 ZTestsPerCaseOct+
                 ZGDP.capita+
                 ZGini+
                 Zlog_PopDensity+
                 ZpercentMigrants+
                 ZGovernment.efficiency+
                 ZTightness+
                 ZStringencyIndex, data=data))

```

### Creating table with all 10 models

```{r model10, echo=FALSE}

#using the tab_model function to inlcude all models in one table

tab_model(Model1, Model2, Model3, Model4, Model5, Model6, Model7, Model8, Model9, Model10,dv.labels = c('Model 1', 'Model 2', 'Model 3', 'Model 4', 'Model 5', 'Model 6', 'Model 7', 'Model 8', 'Model 9', 'Model 10'))

```

## Reproducing Table 2

### Logged deaths per million (7 regression models with different covariats)

```{r table2, echo=FALSE}

#creating the regressions with the variables included in table 2.
#using the lm function and summarizing them by using summary

summary(Model1<-lm(OCTDeaths~
                      ZMortality+
                      ZGDP.capita+
                      ZGini+
                      Zlog_PopDensity+
                      ZpercentMigrants+
                      ZGovernment.efficiency ,data=data))

summary(Model2<-lm(OCTDeaths~
                      ZMortality+
                      ZGDP.capita+
                      ZGini+
                      Zlog_PopDensity+
                      ZpercentMigrants+
                      ZGovernment.efficiency+
                      ZTightness,data=data))

summary(Model3<-lm(OCTDeaths~
                      ZMortality+
                      ZGDP.capita+
                      ZGini+
                      Zlog_PopDensity+
                      ZpercentMigrants+
                      ZGovernment.efficiency+
                      ZTightness+
                      ZCollectivism+
                      ZPowerdistance,data=data))

summary(Model4<-lm(OCTDeaths~
                      ZMortality+
                      ZGDP.capita+
                      ZGini+
                      Zlog_PopDensity+
                      ZpercentMigrants+
                      ZGovernment.efficiency+
                      ZTightness+
                      ZAuthoritarian,data=data))

summary(Model5<-lm(OCTDeaths~
                      ZMortality+
                      ZGDP.capita+
                      ZGini+
                      Zlog_PopDensity+
                      ZpercentMigrants+
                      ZGovernment.efficiency+
                      ZTightness+
                      ZMedian_Age,data=data))

summary(Model6<-lm(OCTDeaths~
                      ZMortality+
                      ZGDP.capita+
                      ZGini+
                      Zlog_PopDensity+
                      ZpercentMigrants+
                      ZGovernment.efficiency+
                      ZTightness+
                      Zday_diff,data=data))

summary(Model7<-lm(OCTDeaths~
                      ZMortality+
                      ZGDP.capita+
                      ZGini+
                      Zlog_PopDensity+
                      ZpercentMigrants+
                      ZGovernment.efficiency+
                      ZTightness+
                      ZStringencyIndex,data=data))

```

### Creating table with all 7 models

```{r model7, echo=FALSE}

#using the tab_model function to inlcude all models in one table


tab_model(Model1, Model2, Model3, Model4, Model5, Model6, Model7,dv.labels = c('Model 1', 'Model 2', 'Model 3', 'Model 4', 'Model 5', 'Model 6', 'Model 7'))


```


## Discussion and reflection

Overall, my data replication project was very successful and I was able to replicate the models as well as the plots. Interestingly, the open-access code for the figures didn't actually produce the plot they had in the paper, but I was able to use some of the ggplot features and create an almost identical graph. 

Otherwise, I found the replication project relatively easy, since the data was already cleaned and well organized, which I personally find the hardest part of data analysis. 





